Capstone Movielens Report

### 2025-11-16 reports/capstone-movielens.site/05.appendix.Rmd
CReport:, function described: tune.model_param.

#### [tune.model_param](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L116) Function {#func.tune.model_param}
\

The function searches for the parameter value corresponding to the minimum value of the `RMSE` from the list of values specified by the `param_values` parameter.

@@

  - **param_values:** A list of values to search for the value corresponding to the minimum value of the `RMSE` ;

::: { #func.tune.model_param.args.fn_tune.test.param_value .sidebar }
  - **fn_tune.test.param_value:** A helper function that calculates the value of the `RMSE` for a given parameter value.;

:::  
  - **break.if_min = TRUE:** A Boolean parameter that determines whether the function should terminate after completing the number of steps specified by the parameter `steps.beyond_min`, after the minimum value of the `RMSE` has been found;

  - **steps.beyond_min = 2:** (takes effect only if `break.if_min` parameter is `TRUE`) Specifies the number of steps after finding the minimum value of the `RMSE`, upon completion of which the function should terminate.

##### Details
\
During execution, the function uses a helper function specified by the `fn_tune.test.param_value` parameter, which calculates the `RMSE` value for the given parameter from the list determined by the `param_values` parameter.

::: {.noteblock data-latex=""}
The algorithm assumes that the dependence of the `RMSE` on the input parameter is a monotonically decreasing function until a minimum is reached and monotonically increasing thereafter. That is, it is assumed that the function has a single minimum on the given interval.
:::

::: { #func.tune.model_param.value .sidebar }
##### Value
\
The function returns a data structure containing the found value of the input parameter `param_values` for which the `RMSE` value is minimal, as well as the minimum `RMSE` value itself, along with a sequence of all calculated `RMSE` values:
```{r eval=FALSE}
  list(tuned.result = data.frame(RMSE = RMSEs_tmp,
                                parameter.value = param_vals_tmp),
       best_result = param_values.best_result)
```
:::

##### Source Code
\
Below is the most significant part of the source code of the [tune.model_param](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L116) function:
```{r eval=FALSE}
tune.model_param <- function(param_values, 
                             fn_tune.test.param_value, 
                             break.if_min = TRUE,
                             steps.beyond_min = 2){
  n <- length(param_values)
  param_vals_tmp <- numeric()
  RMSEs_tmp <- numeric()
  RMSE_min <- Inf
  i_max.beyond_RMSE_min <- Inf
  prm_val.best <- NA

@@

>>>>

#### [tune.model_param](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L126) Function {#func.tune.model_param}
\

Implements an algorithm to find a value of the *regularization parameter* $\lambda$ that corresponds to the minimum value of the *Root Mean Squared Error (RMSE)* for the model under test (we call this the *best value of the regularization parameter* or *best value of $\lambda$*) from the sorted *parameter vector* passed in the [param_values](#func.tune.model_param.args.param_values) argument.

@@

::: { #func.tune.model_param.args .sidebar }
##### Arguments
\

::: { #func.tune.model_param.args.param_values .sidebar }
  - **param_values:** Vector containing a sorted list of $\lambda$ values to search for the best one during the model regularization process;

::: { #func.tune.model_param.args.fn_tune.test.param_value .sidebar }
  - **fn_tune.test.param_value:** Model-specific helper function that validates the model by computing the model's *RMSE* value for a given value of the $\lambda$ parameter;

::: { #func.tune.model_param.args.break.if_min .sidebar }
  - **break.if_min:** (`TRUE` by default) Boolean value that determines whether the algorithm should terminate after completing the number of steps specified by the [steps.beyond_min](#func.tune.model_param.args.steps.beyond_min) argument from the moment the minimum `RMSE` value is found;

::: { #func.tune.model_param.args.steps.beyond_min .sidebar }
  - **steps.beyond_min:** (default 2; only takes effect if [break.if_min](func.tune.model_param.args.break.if_min) is `TRUE`) Specifies the number of steps after finding the minimum *RMSE* value after which the algorithm should terminate.

##### Details
\
The algorithm iterates through the $\lambda$ parameter values from the list passed in the [param_values](#func.tune.model_param.args.param_values) argument, computing the model's *RMSE* for each of them, using the helper function specified by the [fn_tune.test.param_value](#fn_tune.test.param_value) argument.

When the minimum *RMSE* value is reached, the algorithm terminates if [break.if_min](#func.tune.model_param.args.break.if_min) is `FALSE`.
Otherwise, the algorithm continues iterating and terminates after the number of cycles specified by the [steps.beyond_min](#func.tune.model_param.args.steps.beyond_min) argument.

::: {.noteblock data-latex=""}
The algorithm assumes that the dependence of *RMSE* on the parameter $\lambda$ is a monotonically decreasing function until a minimum is reached and monotonically increasing thereafter. That is, it is assumed that the function can have only one minimum on a given interval.
:::

::: { #func.tune.model_param.value .sidebar }
##### Value
\

This function returns a list of two objects:

  - A vector containing a *pair of values*: the found *best value of the regularization parameter* $\lambda$ for which the *RMSE* value is minimal, along with the minimal *RMSE* value itself;
```{r eval=FALSE}
  param_values.best_result <- c(param.best_value, 
                                best_RMSE)
```

  - A data frame containing two variables:

    - *parameter.value*: A vector containing all processed values of the parameter $\lambda$;

    - *RMSE*: A vector containing the *RMSE* values corresponding to the processed values of $\lambda$.
```{r eval=FALSE}
tuned.result = data.frame(RMSE, parameter.value)
```
:::

##### Source Code
\
Below is the most significant part of the source code of the [tune.model_param](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L126) function:
```{r eval=FALSE}
tune.model_param <- function(param_values, 
                             fn_tune.test.param_value, 
                             break.if_min = TRUE,
                             steps.beyond_min = 2){
  n <- length(param_values)
  param_vals_tmp <- numeric()
  RMSEs_tmp <- numeric()
  RMSE_min <- Inf
  i_max.beyond_RMSE_min <- Inf
  prm_val.best <- NA

@@

::: {.noteblock data-latex=""}
The source code of the [tune.model_param](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L126) function is also available in the [Regularization](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L69) section of the [common-helper.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L126) script on _GitHub_.
:::
